---
title: "Reproducible Research: Peer Assessment 1"
output: 
    html_document:
        keep_md: true
---


```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```


## Loading and preprocessing the data
```{r}
# read data
if (!file.exists("activity.csv")) {
  unzip("activity.zip")
}


data<- read.csv("activity.csv")

summary(data)

#Remove any NA's within the Data
data2 <- na.omit(data)

#Change date column into date format
data2$date <- as.Date(data2$date)


summary(data2)
```


## What is mean total number of steps taken per day?

```{r}

data3 <- aggregate(steps ~ date, data2, sum)
hist(data3$steps
     ,breaks = "Sturges"
     , xlab = "Number of Steps"
     , main = "Histogram of Daily Number of Steps"
     ,col = "light blue")
    
mean(data3$steps)
median(data3$steps)

```

The mean number of steps per day = `r round(mean(data3$steps))`

The median number of steps per day = `r median(data3$steps)`


## What is the average daily activity pattern?

```{r}
#The average number of steps per interval
stepsbyint <- aggregate(steps ~ interval, data2, mean)

plot(stepsbyint$interval
     , stepsbyint$steps
     , type ="l"
     , col = "dark blue"
     , ylab = "Number of Steps"
     , xlab = "Interval"
     , main = "Average Number of Steps by Interval")

stepsbyint$interval[which.max(stepsbyint$steps)]
round(stepsbyint$steps[which.max(stepsbyint$steps)])


```
The interval which contains the max number of steps = `r stepsbyint$interval[which.max(stepsbyint$steps)]` with `r round(stepsbyint$steps[which.max(stepsbyint$steps)])` steps

## Imputing missing values

```{r}
sum(is.na(data))

```
The number of NA's in the original data set = `r sum(is.na(data))`

Replacing the NA's with the mean number of steps

```{r}
data4 <- data
data4$date <- as.Date(data4$date)
#Replace the NA's with mean number of steps
data4$steps[is.na(data4$steps)] <- mean(stepsbyint$steps)

summary(data4)
summary(data)

```

Comparing the summaries, it can be seen that the NA's have been removed

Histogram with imputed values:

```{r}
data5 <- aggregate(steps ~ date, data4, sum)
hist(data5$steps
     ,breaks = "Sturges"
     , xlab = "Number of Steps"
     , main = "Histogram of Daily Number of Steps"
     ,col = "light blue")

```

## Are there differences in activity patterns between weekdays and weekends?

Separating the data into weekends and weekdays:

```{r}

day <- weekdays(data4$date)
daytype <- ifelse(day == "Saturday" | day == "Sunday", "Weekend", "Weekday")
data6 <- cbind(daytype, data4)

#Plot to show weekday vs weekend steps

stepsbydaytype <- aggregate(steps ~ interval + daytype, data6, mean)

library(ggplot2)

ggplot(stepsbydaytype, aes(interval, steps)) +
    geom_line(stat = "identity", aes(colour = daytype)) +
    facet_grid(daytype ~ ., scales="fixed", space="fixed")
        
     
     

```
